---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# usa

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/usa)](https://CRAN.R-project.org/package=usa)
[![Travis build status](https://travis-ci.org/kiernann/usa.svg?branch=master)](https://travis-ci.org/kiernann/usa)
[![Codecov test coverage](https://codecov.io/gh/kiernann/usa/branch/master/graph/badge.svg)](https://codecov.io/gh/kiernann/usa?branch=master)
<!-- badges: end -->

The goal of `usa` is to provide updated versions of the `state` data sets
included in base R. These versions include newer statistics in the modern
data frame format.

## Installation

You can install the development version of `usa` from [GitHub](https://github.com) with:

``` r
# install.packages("remotes")
remotes::install_github("kiernann/usa")
```

## Old Data

```{r libs, warning=FALSE, message=FALSE, error=FALSE}
library(tidyverse)
```

R ships with seven outdated built-in "state" objects:

* `state.abb`: character vector of 2-letter abbreviations for the state names.
* `state.area`: numeric vector of state areas (in square miles).
* `state.center`: list giving the approximate geographic center of each state.
* `state.division`: factor giving state divisions.
* `state.name`: character vector giving the full state names.
* `state.region`: factor giving the region.
* `state.x77`: matrix with 50 rows and 8 columns:
  * `Population`: population estimate as of July 1, 1975
  * `Income`: per capita income (1974)
  * `Illiteracy`: illiteracy (1970, percent of population)
  * `Life Exp`: life expectancy in years (1969–71)
  * `Murder`: murder and non-negligent manslaughter rate per 100,000 population (1976)
  * `HS Grad`: percent high-school graduates (1970)
  * `Frost`: mean number of days with minimum temperature below freezing (1931–1960)
  * `Area`: land area in square miles

```{r vector_tibble}
library(datasets)
vectors <- data.frame(
  abb = state.abb,
  name = state.name,
  region = state.region,
  division = state.division,
  x = state.center$x,
  y = state.center$y
)
head(vectors, 3)
head(state.x77, 3)
```

## New Data

This package contains new versions of these objects!

```{r join_states, message=FALSE}
library(usa)
states %>%
  inner_join(abbs) %>%
  inner_join(fips) %>%
  left_join(regions) %>%
  left_join(divisions) %>%
  inner_join(ansi) %>%
  select(-ends_with("id"))
```


***

Please note that the `usa` project is released with a 
[Contributor Code of Conduct](https://kiernann.com/usa/CODE_OF_CONDUCT.html). By
contributing to this project, you agree to abide by its terms.
